NanoMaLy

### 2015-11-18 paper/pldi16-submission/semantics.tex
narrow should take the current substitution.
it needs to apply the su to a hole if possible, to preserve sharing

\paragraph{Narrowing Types} The procedure $\force{e}{t}$ takes as
input an term \texttt{e} and refines it to have the type \texttt{t}
by yielding a pair of either the same expression and a value substitution
the refines the holes in \texttt{e} to ensure it has type \texttt{t},
or yields the stuck term if no such refinement is possible:
%
$$
\begin{array}{lcl}
% \multicolumn{3}{l}{\forcesym \ :: \  (e, t) \rightarrow \pair{e}{\vsu}} \\
% \forcesym            & ::     & (e, t) \rightarrow \pair{e}{\vsu} \\
\force{\ehole{i}}{t} & \defeq & \elet{v}{\gen{t}}{\pair{v}{\ehole{i} \mapsto v}} \\
\force{v}{\ehole{}}  & \defeq & \pair{v}{\emptysu} \\
\force{n}{\tint}     & \defeq & \pair{n}{\emptysu} \\
\force{v}{\tint}     & \defeq & \pair{\stuck}{\emptysu} \\
\force{b}{\tbool}    & \defeq & \pair{b}{\emptysu} \\
\force{v}{\tbool}    & \defeq & \pair{\stuck}{\emptysu} \\
\force{\efun{x}{e}}{\tfun{\thole{}}{\thole{}}} & \defeq & \pair{\efun{x}{e}}{\emptysu} \\
\force{v}{\tfun{\thole{}}{\thole{}}} & \defeq & \pair{\stuck}{\emptysu} \\
\end{array}
$$

@@

$$
\begin{array}{lcll}
% \gensym       & ::      & t \rightarrow v \\
\gen{\tint}   & \defeq  & n & \quad \text{non-deterministic} \\
\gen{\tbool}  & \defeq  & b & \quad \text{non-deterministic} \\
\gen{\tfun{t_1}{t_2}}   & \defeq & \efun{x}{\ehole{i}}, & \quad \text{$i$ is fresh} \\
\gen{\thole{}} & \defeq & \ehole{i}, & \quad \text{$i$ is fresh} \\
\end{array}
$$

@@

\begin{thm}{\textbf{[Witness Generality]}}
\label{thm:soundness}
  For any function $f$, if \steptr{\trace}{\eapp{f}{\ehole{}}}{\emptysu}{\stuck}{\su},
  then for every type $t$ there exists $\hastype{v}{t}$ such that
  \steps{\eapp{f}{v}}{\emptysu}{\stuck}{\su}$.
\end{thm}

@@

\paragraph{Modeling Semantics}
%: \texttt{eval} and \texttt{isStuck}}
%
We resolve the non-determinism in the operational semantics
(\S~\ref{sec:semantics}) via the procedure:
%
\begin{mcode}
  eval :: ($e$, $\su$) -> [($v$, $\su$)]
\end{mcode}
%
Due to the non-determinism introduced by \gensym, a call @eval($e$, $\vsu$)@
returns the \emph{list} possible results of the form $\pair{e'}{\vsu'}$ such
that $\steps{e}{\vsu}{e'}{\vsu'}$.
%
We determine whether a given value is stuck by using the predicate
%
\begin{mcode}
  isStuck :: $v$ -> Bool
\end{mcode}

\paragraph{Generating Target Arguments}
%
We address the issue of currying by defining a procedure @genArgs@ such that
@genArgs [(f1,e1),...,(fn, en)]@ yields a list of holes @[v1,...,vk]@
such that @fn v1 ... vk@ \emph{does not} evaluate to a lambda.
%
This is achieved via a simple @loop@ that keeps adding holes until
@eval@uating the closed term yields a non-lambda value.
%
The helper functions @witness@ and @close@ are used to respectively
apply the parameters to the target and close the result under the top-level
let-binders, prior to invoking @eval@.
%
That is, @mkApps@ creates a nested sequence of applications in
the usual left-associative style, and @mkLets@ takes a list of
binders and a body expression, and creates a sequence of nested
let-binders that close the body expression.

@@

\begin{thm}{\textbf{[Witness Generation]}}
\label{thm:generation}
  If @genWitness k@ $[(f_1, e_1), \ldots, (f_n, e_n)]$ is non-empty,
  then for every type $t$ there exists $\hastype{v}{t}$ such that
  \steps{\eapp{f_n}{v}}{\emptysu}{\stuck}{\su}$.
\end{thm}

>>>>

\paragraph{Narrowing Types} The procedure $\force{e}{t}{\vsu}$ takes as
input an term \texttt{e} and refines it to have the type \texttt{t}
by yielding a pair of either the same expression and a value substitution
the refines the holes in \texttt{e} to ensure it has type \texttt{t},
or yields the stuck term if no such refinement is possible:
%
$$
\begin{array}{lcl}
% \multicolumn{3}{l}{\forcesym \ :: \  (e, t) \rightarrow \pair{e}{\vsu}} \\
% \forcesym            & ::     & (e, t) \rightarrow \pair{e}{\vsu} \\
\force{\ehole{i}}{t}{\vsu} & \defeq & \hspace{-1ex} \begin{array}{l}
    \mathtt{if}\ i \in \vsu\ \mathtt{then}\ \pair{\vsu(i)}{\emptysu}\ \mathtt{else} \\
         \elet{v}{\gen{t}}{\pair{v}{\ehole{i} \mapsto v}}
  \end{array} \\
\force{v}{\ehole{}}{\vsu}  & \defeq & \pair{v}{\emptysu} \\
\force{n}{\tint}{\vsu}     & \defeq & \pair{n}{\emptysu} \\
\force{v}{\tint}{\vsu}     & \defeq & \pair{\stuck}{\emptysu} \\
\force{b}{\tbool}{\vsu}    & \defeq & \pair{b}{\emptysu} \\
\force{v}{\tbool}{\vsu}    & \defeq & \pair{\stuck}{\emptysu} \\
\force{\efun{x}{e}}{\tfun{\thole{}}{\thole{}}}{\vsu} & \defeq & \pair{\efun{x}{e}}{\emptysu} \\
\force{v}{\tfun{\thole{}}{\thole{}}}{\vsu} & \defeq & \pair{\stuck}{\emptysu} \\
\end{array}
$$

@@

$$
\begin{array}{lcll}
% \gensym       & ::      & t \rightarrow v \\
\gen{\tint}   & \defeq  & n, & \quad \text{non-deterministic} \\
\gen{\tbool}  & \defeq  & b, & \quad \text{non-deterministic} \\
\gen{\tfun{t_1}{t_2}}   & \defeq & \efun{x}{\ehole{i}}, & \quad \text{$i$ is fresh} \\
\gen{\thole{}} & \defeq & \ehole{i}, & \quad \text{$i$ is fresh} \\
\end{array}
$$

@@

\begin{thm}{\textbf{[Witness Generality]}}
\label{thm:soundness}
  For any function $f$, if \steptr{\trace}{\eapp{f}{\ehole{}}}{\emptysu}{\stuck}{\su},
  then for every type $t$ there exists $\hastype{v}{t}$ such that
  $\steps{\eapp{f}{v}}{\emptysu}{\stuck}{\su}$.
\end{thm}

@@

\paragraph{Modeling Semantics}
%: \texttt{eval} and \texttt{isStuck}}
%
We resolve the non-determinism in the operational semantics
(\S~\ref{sec:semantics}) via the procedure:
%
\begin{mcode}
  eval :: ($e$, $\su$) -> [($v$, $\su$)]
\end{mcode}
%
Due to the non-determinism introduced by \gensym, a call \hbox{|eval($e$, $\vsu$)|}
returns the \emph{list} possible results of the form $\pair{e'}{\vsu'}$ such
that $\steps{e}{\vsu}{e'}{\vsu'}$.
%
We determine whether a given value is stuck by using the predicate
%
\begin{mcode}
  isStuck :: $v$ -> Bool
\end{mcode}

\paragraph{Generating Target Arguments}
%
We address the issue of currying by defining a procedure @genArgs@ such that \\
\hbox{@genArgs [(f1,e1),...,(fn,en)]@} yields a list of holes @[v1,...,vk]@
such that \hbox{@fn v1 ... vk@} \emph{does not} evaluate to a lambda.
%
This is achieved via a simple @loop@ that keeps adding holes until
@eval@uating the closed term yields a non-lambda value.
%
The helper functions @witness@ and @close@ are used to respectively
apply the parameters to the target and close the result under the top-level
let-binders, prior to invoking @eval@.
%
That is, @mkApps@ creates a nested sequence of applications in
the usual left-associative style, and @mkLets@ takes a list of
binders and a body expression, and creates a sequence of nested
let-binders that close the body expression.

@@

\begin{thm}{\textbf{[Witness Generation]}}
\label{thm:generation}
  If @genWitness k@ $[(f_1, e_1), \ldots, (f_n, e_n)]$ is non-empty,
  then for every type $t$ there exists $\hastype{v}{t}$ such that
  $\steps{\eapp{f_n}{v}}{\emptysu}{\stuck}{\su}$.
\end{thm}
